---
title: "Work Sheet #2"
author: "Mikee Lorraine S. Benjamin"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1 – Create a vector using : operator

### a. Sequence from -5 to 5  
```{r}
y <- -5:5
y
```

### b. x <- 1:7  
```{r}
x <- 1:7
x
```

---

## Exercise 2 – Create a vector using seq()

### a. seq(1, 3, by=0.2)  
```{r}
vec <- seq(1, 3, by = 0.2)
vec
```

---

## Exercise 3 – Worker Ages

```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41,
51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30,
61, 54, 58, 26, 18)
```

### a. Access 3rd element  
```{r}
ages[3]
```

### b. Access 2nd and 4th  
```{r}
ages[c(2,4)]
```

### c. Access all except the 1st  
```{r}
ages[-1]
```

---

## Exercise 4 – Named Vector

```{r}
x <- c(first = 3, second = 0, third = 9)
x
```

### Access `"first"` and `"third"`
```{r}
x[c("first", "third")]
```

---

## Exercise 5 – Modify a Vector

```{r}
x <- -3:2
x
```

### Replace 2nd element with 0
```{r}
x[2] <- 0
x
```

---

## Exercise 6 – Fuel Data

### a. Create data frame
```{r}
month <- c("Jan", "Feb", "March", "Apr", "May", "June")
price <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
quantity <- c(25, 30, 40, 50, 10, 45)

fuel_data <- data.frame(Month = month, Price = price, Quantity = quantity)
fuel_data
```

### b. Average expenditure (weighted mean)
```{r}
avg_expenditure <- weighted.mean(price, quantity)
avg_expenditure
```

---

## Exercise 7 – Rivers Dataset Analysis

### a. Build vector of statistics  
```{r}
data <- c(
  length(rivers),
  sum(rivers),
  mean(rivers),
  median(rivers),
  var(rivers),
  sd(rivers),
  min(rivers),
  max(rivers)
)

names(data) <- c("Length", "Sum", "Mean", "Median", "Variance", "SD", "Min", "Max")
data
```

---

## Exercise 8 – Forbes Celebrity Data

### a. Create vectors  
```{r}
power_rank <- 1:25

celeb_names <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2",
"Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent",
"Cast of Sopranos", "Dan Brown", "Bruce Springsteen", "Donald Trump",
"Muhammad Ali", "Phaul McCartney", "George Lucas", "Elton John",
"David Letterman", "Phil Mickelson", "J.K Rowling", "Brad Pitt",
"Peter Jackson", "Dr. Phil McGraw", "Jay Lenon", "Celine Dion",
"Kobe Bryant")

pay <- c(65, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40,
         233, 44, 40, 47, 75, 25, 39, 45, 32, 40, 31)

forbes <- data.frame(
  Power_Rank = power_rank,
  Celebrity_Name = celeb_names,
  Pay = pay
)

forbes
```

### b. Modify J.K. Rowling’s ranking + pay  
```{r}
idx <- which(forbes$Celebrity_Name == "J.K Rowling")
forbes$Power_Rank[idx] <- 15
forbes$Pay[idx] <- 90

forbes[idx, ]
```

### c. Interpretation  
J.K. Rowling’s updated rank of **15** and higher pay reflect her strong cultural and financial influence.

