---
title: "Work Sheet #3a"
author: "Mikee Lorraine S. Benjamin"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

There is a built-in vector **LETTERS** (uppercase letters) and **letters** (lowercase letters).

## LETTERS:

### a. Produce a vector that contains the first 11 letters.
```{r}
LETTERS[1:11]
```

### b. Produce a vector that contains the odd-numbered letters.
```{r}
LETTERS[seq(1, 26, by = 2)]
```

### c. Produce a vector that contains the vowels.
```{r}
LETTERS[LETTERS %in% c("A", "E", "I", "O", "U")]
```

## letters:

### d. Produce a vector that contains the last 5 lowercase letters.
```{r}
tail(letters, 5)
```

### e. Produce a vector that contains lowercase letters 15 to 24.
```{r}
letters[15:24]
```

---

# Exercise 2

Create a vector with the average April temperatures in Tuguegarao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City.

### a. Create a character vector for the cities.
```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban",
          "Samal Island", "Davao City")
city
```

### b. Create a numeric vector of temperatures.
```{r}
temp <- c(42, 39, 34, 34, 30, 27)
temp
```

### c. Create a dataframe combining city and temp.
```{r}
weather <- data.frame(city, temp)
weather
```

### d. Rename the dataframe columns to City and Temperature.
```{r}
names(weather) <- c("City", "Temperature")
weather
```

### e. Print the structure of the dataframe.
```{r}
str(weather)
```

### f. Display rows 3 and 4.
```{r}
weather[3:4, ]
```

### g. Display the city with highest and lowest temperature.
```{r}
weather[which.max(weather$Temperature), ]
weather[which.min(weather$Temperature), ]
```

---

# Exercise 3  
Create a matrix of 1–8 and 11–14 with 4 columns and 3 rows.

### a. Create the matrix.
```{r}
m <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
m
```

### b. Multiply the matrix by 2.
```{r}
m2 <- m * 2
m2
```

### c. Display row 2.
```{r}
m2[2, ]
```

### d. Display column 3 and 4 of rows 1 and 2.
```{r}
m2[1:2, 3:4]
```

### e. Display columns 2 and 3 of row 3.
```{r}
m2[3, 2:3]
```

### f. Display only column 4.
```{r}
m2[, 4]
```

### g. Name the rows and columns.
```{r}
rownames(m2) <- c("isa", "dalawa", "tatlo")
colnames(m2) <- c("uno", "dos", "tres", "quatro")
m2
```

### h. Reshape the original matrix into 6 rows and 2 columns.
```{r}
dim(m) <- c(6, 2)
m
```

---

# Exercise 4  
An array contains: 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1 (repeat twice)

### a. Create a 3D array with 2 rows, 4 columns, and 3 layers.
```{r}
values <- rep(c(1,2,3,6,7,8,9,0,3,4,5,1), times = 2)
arr <- array(values, dim = c(2, 4, 3))
arr
```

### b. How many dimensions?
```{r}
dim(arr)
```

### c. Add row, column, and layer names.
```{r}
dimnames(arr) <- list(
  letters[1:2],
  LETTERS[1:4],
  c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array")
)
arr
```
