---
title: "Rworksheet_benjamin#3b"
author: "Mikee Lorraine S. Benjamin"
date: "2025-10-20"
output:
  pdf_document:
    latex_engine: xelatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Create a data frame using the table below.

## a. Write the codes.
```{r}
# Create the data frame
df <- data.frame(
  Respondents = 1:20,
  Sex = c(2,2,1,2,2,2,2,1,2,2,1,2,2,2,2,2,2,2,1,2),
  Fathers_Occupation = c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1),
  Persons_at_Home = c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6),
  Siblings_at_School = c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,2,2,5,3,2),
  Types_of_Houses = c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)
)
df
```

## b. Describe the data.
```{r}
str(df)
summary(df)
```

## c. Is the mean number of siblings attending school equal to 5?
```{r}
mean(df$Siblings_at_School)
```

## d. Extract the 1st two rows.
```{r}
df[1:2, ]
```

## e. Extract 3rd and 5th row with 2nd and 4th column.
```{r}
df[c(3,5), c(2,4)]
```

## f. Select the variable Types_of_Houses into a vector.
```{r}
types_houses <- df$Types_of_Houses
types_houses
```

## g. Select all male respondents whose father's occupation was Farmer.
```{r}
subset(df, Sex == 1 & Fathers_Occupation == 1)
```

## h. Select all female respondents with ≥5 siblings attending school.
```{r}
subset(df, Sex == 2 & Siblings_at_School >= 5)
```

---

# 2. Create an empty data frame

```{r}
df_empty <- data.frame(
  Ints = integer(),
  Doubles = double(),
  Characters = character(),
  Logicals = logical(),
  Factors = factor(),
  stringsAsFactors = FALSE
)

str(df_empty)
```

---

# 3. Create a CSV file then import it

## a. Import the CSV file

First create the object to save:

```{r}
household <- df
write.csv(household, "HouseholdData.csv", row.names = FALSE)

data <- read.csv("HouseholdData.csv")
data
```

## b. Convert Sex to factor and then integer  
(Male = 1, Female = 2)

```{r}
data$Sex <- factor(data$Sex, levels = c(1, 2), labels = c("Male", "Female"))
data$Sex <- as.integer(data$Sex)
data
```

## c. Convert Type of House into factor (Wood=1; Concrete=2; Semi-Concrete=3)

```{r}
data$Types_of_Houses <- factor(
  data$Types_of_Houses,
  levels = c(1,2,3),
  labels = c("Wood", "Concrete", "Semi-Concrete")
)

data$Types_of_Houses <- as.integer(data$Types_of_Houses)
data$Types_of_Houses
```

## d. Factor father's occupation (Farmer=1; Driver=2; Others=3)
```{r}
data$Fathers_Occupation <- factor(
  data$Fathers_Occupation,
  levels = c(1,2,3),
  labels = c("Farmer", "Driver", "Others")
)

data$Fathers_Occupation <- as.integer(data$Fathers_Occupation)
data
```

## e. Select female respondents whose father is a driver.
```{r}
female_driver <- subset(data, Sex == 2 & Fathers_Occupation == 2)
female_driver
```

## f. Select respondents with >= 5 siblings attending school.
```{r}
siblings5 <- subset(data, Siblings_at_School >= 5)
siblings5
```

---

# 4. Interpretation of the Graph

The chart indicates that most respondents live in wooden houses, followed by concrete and semi-concrete types.  
Those living in wooden houses tend to have more siblings attending school, suggesting larger families.  
In contrast, residents of concrete houses generally have fewer siblings, implying smaller family size and possibly higher economic stability.

