---
title: "Work Sheet 4a"
author: "Mikee Lorraine S. Benjamin"
date: "2025-11-17"
output:
  pdf_document:
    latex_engine: xelatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data Frame Creation

```{r}
household <- data.frame(
  ShoeSize = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 
               10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 
               7.5, 10.5, 8.5, 11.0, 9.0, 13.0),
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 73.0, 72.0, 64.0,
             74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 
             64.0, 73.0, 69.0, 70.0, 69.0, 70.0),
  Gender = c("F", "F", "F", "F", "F", "F", "F", "M", "M", "F",
             "F", "F", "M", "M", "M", "M", "F", "F", "F", "F",
             "F", "M", "M", "M", "M", "M")
)

household
```

### b. Subset by gender
```{r}
male <- subset(household, Gender == "M", select = c(ShoeSize, Height))
female <- subset(household, Gender == "F", select = c(ShoeSize, Height))

male
female
```

### c. Mean of shoe size and height
```{r}
mean_shoesize <- mean(household$ShoeSize)
mean_height <- mean(household$Height)

mean_shoesize
mean_height
```

### d. Relationship  
There is a positive relationship: taller individuals tend to have larger shoe sizes.

---

## 2. Character to Factor
```{r}
months <- c("March","April","January","November","January","September","October",
            "September","November","August","January","November","November",
            "February","May","August","July","December","August","August",
            "September","November","February","April")

factor_months_vector <- factor(months)
factor_months_vector
```

---

## 3. Summary of character vs factor
```{r}
summary(months)
summary(factor_months_vector)
```

---

## 4. Ordered Factor
```{r}
direction <- c("East", "West", "West", "West", "West", "North", "North", "North")
factor_data <- factor(direction, levels = c("East", "West", "North"))
factor_data
summary(factor_data)
```

---

## 5. Importing Data

### a. Import Excel file  
Make sure `import_march.xlsx` is in your working directory.

```{r}
library(readxl)
import_march <- read_excel("import_march.xlsx")
import_march
```

### If your teacher requires CSV format instead:
```{r, eval=FALSE}
import_march <- read.table("import_march.csv", header = TRUE, sep = ",")
import_march
```

---

## 6. Full Search
```{r}
num <- 15
numbers <- 1:50
found <- FALSE

for (i in numbers) {
  if (i == num) {
    found <- TRUE
    break
  }
}

if (!found) {
  cat("The number selected is beyond the range of 1 to 50\n")
} else if (num == 20) {
  cat("TRUE\n")
} else {
  cat("You selected:", num, "\n")
}
```

---

## 7. Minimum Bills Function
```{r}
minimum_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  remaining <- price
  count <- 0
  
  for (bill in bills) {
    if (remaining >= bill) {
      num <- remaining %/% bill
      count <- count + num
      remaining <- remaining - num * bill
    }
  }
  
  cat("Minimum number of bills needed:", count, "\n")
}

price_of_snack <- 1350
minimum_bills(price_of_snack)
```

---

## 8. Student Grades

### a. Data frame
```{r}
grades <- data.frame(
  Name = c("Annie", "Thea", "Steve", "Hanna"),
  Grade1 = c(85, 65, 75, 95),
  Grade2 = c(65, 75, 55, 75),
  Grade3 = c(85, 90, 80, 100),
  Grade4 = c(100, 90, 85, 90)
)

grades
```

### b. Average ≥ 90 (no rowMeans)
```{r}
for (i in 1:nrow(grades)) {
  total <- grades$Grade1[i] + grades$Grade2[i] +
           grades$Grade3[i] + grades$Grade4[i]
  avg <- total / 4
  if (avg >= 90) {
    cat(grades$Name[i], "'s average grade this semester is", avg, "\n")
  }
}
```

### c. Tests where mean < 80
```{r}
for (i in 2:ncol(grades)) {
  total <- sum(grades[, i])
  avg <- total / nrow(grades)
  
  if (avg < 80) {
    cat("The", i - 1, "th test was difficult.\n")
  }
}
```

### d. Highest score > 90 (no max)
```{r}
for (i in 1:nrow(grades)) {
  highest <- grades[i, 2]
  for (j in 3:ncol(grades)) {
    if (grades[i, j] > highest) {
      highest <- grades[i, j]
    }
  }
  if (highest > 90) {
    cat(grades$Name[i], "'s highest grade this semester is", highest, "\n")
  }
}
```

